\documentclass[12pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{mathbbol}
\usepackage{hyperref}
\geometry{margin=1in}

\title{Report: Max-Cut Problem}
\author{Hanna Makowska}
\date{19 March 2026}

\begin{document}

\maketitle

\section{Problem Statement}
The goal of this task is to compute a max cut in a weighted undirected graph.
Given a graph $G=(V,E)$ with edge weights $w_{uv}$, we want to divide the set of
vertices into two sets $S$ and $\bar{S}$ to maximize the total weight of edges
crossing the partition.

\section{Input Format}
The input file is divided into two parts:
\begin{itemize}
    \item A block starting with \texttt{V}, listing vertices.
    \item A block starting with \texttt{E}, listing weighted edges in the form
          \texttt{(u)-(v)-w}.
\end{itemize}

Example input used in this project:
\begin{verbatim}
V
    (a1)
    (a2)
    (a3)
    (a4)
    (a8)
E
    (a1)-(a4)-40
    (a4)-(a3)-30
    (a8)-(a3)-30
\end{verbatim}

\section{First Approach: Random Partitioning}
The simplest method to find a max cut is to randomly assign each vertex to one of the two sets.
After that we can calculate the total weight of edges crossing the partition, according to the formula:
\[
    max \sum_{(i, j) \in E} w_{ij} x_{i} (1 - x_{j}) + w_{ij} x_{j} (1 - x_{i}),x_{i} \in \{0, 1\}
\]
and repeat this
process multiple times to find the best solution.

\section{Second Approach: Local Search}
A more sophisticated method is to use a local search algorithm. We start with a random partition and
iteratively improve it by moving vertices from one set to the other if it increases the total weight
of crossing edges. The algorithm continues until no single vertex move can improve the solution.

\subsection{Proof}
To prove that the local search algorithm finds a local maximum, we can use the following argument:
Let $S$ be the current partition of vertices into two sets. For each vertex $v$, we calculate the change
in the total weight of crossing edges if we move $v$ from its current set to the other set. If there
exists a vertex $v$ such that moving it increases the total weight, we perform the move and update the partition.
We repeat this process until no vertex can be moved to increase the total weight.
At this point, we have reached a local maximum because any single vertex move would not improve the solution.

\subsection{Complexity}
In worse case scenario we will have to check all vertices and edges in each iteration, which gives us a time complexity
of $O(|V| \cdot |E|)$. However, in practice, the algorithm often converges much faster, especially for sparse graphs.

\section{Reference}
\url{https://www.cs.cmu.edu/afs/cs/academic/class/15854-f05/www/scribe/lec02.pdf}

\end{document}